\chapter{Terminology}
\label{ch:mdl}
\section{Model driven engineering}
Model driven engineering (MDE from now on) is a promising approach to problem solving, especially for complex applications. MDE alleviate the complexity of the problem by using a domain specific language to express domain specific concepts. This aims to maximize compatibility between systems, especially those implemented by different individuals or different teams: such goal is archived by making terminology and design patterns uniform.
While MDE is quite popular in software engineering (especially in large software projects), it is still considered underestimated in the robotics field.

\section{Metamodelling}
A model is an abstraction of a real system and it exactly describes its behavior; multiple systems can conform to the same model, meaning that they behave the same (e.g. multiple implementations of a control system).
The same way, a meta-model is a model of model, meaning that it is yet another abstraction; it is used to describe at a more conceptual level rules and constraints of a given class of models. To make an example, in the study case presented in this thesis the extended Kalman filter conforms to the mathematical model of an extended Kalman filter which in turns conforms to the meta-model of a \emph{Bayesian estimator}, like the models of other types of filters, e.g. particle filters.

\section{Domain specific languages}
Domain specific languages (DSL) are languages specifically designed to easily express schemes or patterns commons within a certain field, without the need of details of some particular implementation. DSL are commonly used to describe models and meta-models. A DSL is not necessarily a new language: it will be show for example that in UBX a particular Lua table is used to describe the structure of a component-based software.
